---
title: "FastLZeroSpikeInference Tutorial"
author: "Sean Jewell"
date: '2018-12-07'
output: html_document
layout: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
require(FastLZeroSpikeInference)
```

R tutorial
----

In this tutorial, we demonstrate how the ```FastLZeroSpikeInference``` package can be used to deconvolve calcium imaging data from a single neuron. All demos require the ```FastLZeroSpikeInference``` package; installation instructions are provided [here](software.html). 

To illustrate the software, we generate a synthetic dataset according to 

$$
y_t = c_t + \epsilon_t, \\
c_t = \gamma c_{t-1} + z_t, \\
z_t \sim \mathrm{Pois}(\theta), \\
$$

where $\epsilon_t \sim \mathrm{N}(0, \sigma^2)$ and  $c_1 = 1$. The function ```simulate_ar1``` generates data from this model. 

```{r}
sim <- simulate_ar1(n = 10000, gam = 0.95, poisMean = 0.01, sd = 0.15, seed = 1)
plot(sim)
```


The ```estimate_spikes``` function estimates spikes based on the calcium trace, exponential decay parameter $\gamma$, and a tuning parameter $\lambda$. This function solves

$$
\underset{c_1,\ldots,c_T, z_2,\ldots,z_T}{\mathrm{minimize}}  
\frac{1}{2} \sum_{t=1}^T \left( y_t -  c_t \right)^2 + \lambda \sum_{t=2}^T 1_{\left( z_t \neq 0 \right) } 
$$

if the ```constraint``` parameter is set to false (default), and 

$$
\underset{c_1,\ldots,c_T, z_2,\ldots,z_T}{\mathrm{minimize}}  
\frac{1}{2} \sum_{t=1}^T \left( y_t -  c_t \right)^2 + \lambda \sum_{t=2}^T 1_{\left( z_t \neq 0 \right) }
\mbox{ subject to } z_t = c_t - \gamma c_{t-1} \geq 0,
$$

if the ```constraint``` parameter is set to true. Here's an example of the function applied to the previously generated data. 


```{r}
# AR(1) model
fit <- estimate_spikes(dat = sim$fl, gam = 0.95, lambda = 1)
print(fit)
```

By default, and to save computation time, the calcium concentration is not automatically estimated. However, the ```estimate_calcium``` function can be used to estimate the calcium concentration based on a prior fit. 

```{r}
# compute fitted values from prev. fit
fit <- estimate_calcium(fit)
plot(fit)
```

Alternatively, both spikes and calcium concentrations can be computed with the ```estimate_spikes``` function with ```estimate_calcium``` parameter set to true. 

```{r}
fit <- estimate_spikes(dat = sim$fl, gam = 0.95, lambda = 1, estimate_calcium = T)
plot(fit)
```

To solve the constrained problem 

$$
\underset{c_1,\ldots,c_T, z_2,\ldots,z_T}{\mathrm{minimize}}  
\frac{1}{2} \sum_{t=1}^T \left( y_t -  c_t \right)^2 + \lambda \sum_{t=2}^T 1_{\left( z_t \neq 0 \right) }
\mbox{ subject to } z_t = c_t - \gamma c_{t-1} \geq 0,
$$

we set the ```constraint``` parameter to true in the ```estimate_spikes``` function. Consider the following example.

```{r}
# Constrained AR(1) model
fit <- estimate_spikes(dat = sim$fl, gam = 0.95, lambda = 1, constraint = T, estimate_calcium = T)
print(fit)
plot(fit)
```



